\section{Formalization}
\label{sec:formalization}

% SpecIR example:
% (bvand a b) (a and b are (bitvector 8))
% LRIR:
% (hash-ref
%  (hw-module-instance 
%   "LUT2"
%   memory = 4'b1000,
%   i0 = a,
%   i1 = b)
%  'O)

% (rosette-synthesize SpecIR-program LRIR-program)
% -> assert SpecIR-program == LRIR-program
% -> solve

% I_S : SpecIR -> bitvector
%  --> "just" Rosette/Racket. (bvand a b) evals automatically.
%  we don't have to do (interpret (bvand a b))
% I_LR : LRIR -> bitvector
%  (lr:hash-ref ...)
%  (interpret (lr:hash-ref ...))

% exists M, 
%  forall inputs, I_S(P_S, inputs, M) 
%                   == I_LR(P_LR, inputs, M).
% synthesis: find M

% in the past/in the current version of the paper:
% I_S and I_LR -> bitvectors
%
% New thing:
% I_S and I_LR -> signals

% This will change the synthesis problem
% Old problem:
% exists M, 
%  forall inputs, I_S(P_S, inputs, M) 
%                   == I_LR(P_LR, inputs, M).
% synthesis: find M

% now we have multiple possible synthesis problems
% exists M, 
%  forall inputs, (signal-value I_S(P_S, inputs, M))
%                   == (signal-value I_LR(P_LR, 
%                                     inputs, M)).
% synthesis: find M

% exists M, 
%  forall inputs, 
%   (let o0 = I_S(P_S, inputs + clk=0, M);
%    let o1 = I_S(P_S, inputs + clk=1 + state(o0), M);
%    let o2 = I_S(P_S, inputs + clk=0 + state(o1), M);
%    ...
%    (signal-value oN))
%   == 
%   (let o0 = I_LR(P_LR, inputs + clk=0, M);
%    let o1 = I_LR(P_LR, inputs + clk=1 + state(o0), M);
%    let o2 = I_LR(P_LR, inputs + clk=0 + state(o1), M);
%    ...
%    (signal-value oN))
% synthesis: find M

In this section, we formalize \lr
  and provide a mathematical argument
  for its correctness.
We start by introducing and 
  formalizing the semantics used in \lr,
  beginning with introducing
  its base value type:
  the \signal datatype (\cref{sec:signal}).
Then, we define syntax and semantics
  of the imported Verilog (\cref{sec:verilog-semantics})
  and of the \lr IR, \lrir (\cref{subsec:spec-ir-and-lrir}).
Finally, we present
  a mathematical model
  of the \lr workflow
  (formalized via the \lrfn function)
  and prove its correctness
  (assuming the correctness of the underlying program synthesis tool and HDL compiler) in \cref{sec:correctness-argument}.
%Then we introduce the \lrfn function
  %to formalize the \lr workflow
  %introduced in \cref{sec:overview}
  %and argue for its correctness
  %via stating the \lr-Correct property
  %(\cref{sec:correctness-argument}).

  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
% Subsection: Signals
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection{The \texttt{signal} Datatype}
\label{sec:signal}

To formalize
  expressions
  in both the
  imported Verilog 
  and \lrir,
  we first introduce their base value type:
  the \signal datatype.
\signal{}s are fundamental to \lr's ability
  to synthesize sequential designs.
While bitvectors alone are expressive enough
  to capture the semantics of 
  \textit{combinational} expressions,
  they are not expressive enough to capture
  \textit{sequential} expressions.
Thus, we introduce \signal{}s
  as a more expressive datatype
  for capturing hardware expressions.
  
A \signal represents
  the value on an $n$-bit
  wire and
  the current state
  of the stateful elements
  in the context where the \signal
  was produced.
Consequently, we represent a
  \signal{} as a pair $(v,m)$,
  where $v$ is a bitvector
  capturing the value on the wire
  and $m$ is a map
  capturing the current state.

% \input{figures/state-merge-and-update}
\input{figures/signal-state}
The state map $m$
  maps keys (which uniquely identify
    an expression's internal state elements)
  to a tuple $(\mathit{sv}, \mathit{vers})$,
  where $\mathit{sv}$
  is the value of the state
  and $\mathit{vers}$ is
  its version number 
  (necessary for handling conflicts during merging).
Beyond indexing into a state map,
  we define two core operations of state maps:
  \textit{merging} and \textit{updating,}
  both defined via pseudocode
  in \cref{fig:signal-complete}.
Merging two state maps
  simply unions the two maps together,
  while handling conflicts by taking the value
  with the highest version number.
Updating a state map
  involves setting a key to a value
  while incrementing the version (or defaulting to 0).
  
  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
% Subsection: Verilog Semantics
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection{Imported Verilog Semantics}
\label{sec:verilog-semantics}

\input{figures/verilog-syntax}

\input{figures/yosys-semantics}

\lr extracts semantics from Verilog,
  both from the user-provided Verilog file
  to be compiled,
  and the vendor-supplied FPGA primtive
  simulation models.
The process for extracting these semantics
  is described in detail in
  \cref{sec:implementation:importing-semantics}.
In this section, we formalize 
  the semantics
  which result from this extraction process.
While this formalization
  is a contribution of this paper,
  the underlying semantics
  of the subset of Verilog
  \lr supports
  are designed and implemented by
  the Yosys synthesis tool~\cite{wolf2013yosys}.

\cref{fig:verilog-syntax}
  lists the syntax for the subset of Verilog 
  supported for import by \lr,
  while \cref{fig:verilog-semantics}
  defines its semantics.
Imported Verilog expressions evaluate to \signal{}s.
In general, the imported Verilog expressions
  cover a wide swath
  of bitvector operations,
  including bitwise operations,
  arithmetic operations,
  comparisons,
  and muxes.

\texttt{dff} represents
  a D flip-flop: a state element
  with a clock and data input,
  which stores the value on the data input
  when the clock changes from low to high.
The semantics for this expression
  are given by \textbf{dff-hold}
  and \textbf{dff-tick}.
\textbf{dff-hold}
  describes 
  the case when the clock doesn't tick,
  i.e., there's no \textit{positive edge}
  going from 0 to 1.
In this case,
  the flip flop expression holds
  its previous value $q'$,
  which it gets from the state map.
\textbf{dff-tick}
  describes the case in which
  the clock ticks,
  and thus the expression
  takes on the new value $d'$.
Note, crucially,
  that the value becomes $d'$
  (the \textit{previous} value of $d$,
  read from the state map)
  and not the current value of $d$.\footnote{
  These semantics are defined by
  Yosys's \texttt{clk2fflogic} pass.
}


The \texttt{dff} expression is the 
  only expression
  in both Verilog and \lrir semantics
  that updates the state map.
All other expressions
  are \textit{combinational},
  meaning they do
  not alter state.
% This is significant as it allows us
%   to simplify our mental model of 
%   imported Verilog and \lrir semantics.
% Knowing that \texttt{dff} is the only
%   expression to update the state map,
%   we can interpret
%   all other expressions
%   in the imported Verilog
%   and \lrir
%   as being purely combinational.
% (Note that expressions generally \textit{merge}
%   state maps, seen in most rules in 
%   \cref{fig:verilog-semantics,fig:lrir-semantics},
%   but crucially, only \texttt{dff} \textit{updates}
%   values in the map).
  


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
% Subsection: Spec IR and LRIR
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection{\lrir}
\label{subsec:spec-ir-and-lrir}

% \lr defines two Intermediate Representations (IRs),
%   \Spec IR and \lrir;
%   syntax for both is defined
%   in \cref{fig:lakeroad-dsl-grammar}.
% \lr uses \Spec IR
%   to give specifications for
%   instructions to synthesize.
% Terms in \Spec IR are either
%   bitvector constants,
%   variables which represent
%   inputs to the instruction,
%   or operators
%   applied to smaller terms.
% \Spec IR is a subset of the SMT-LIB language~\cite{smtlib},
%   from which it derives its semantics.

\input{figures/lakeroad-dsl-grammar}

\input{figures/lakeroad-dsl-semantics}

\lr uses the \lr Intermediate Representation (\lrir) to represent
  wire operations and computations on FPGAs.
The syntax for \lrir is shown in
  \cref{fig:lakeroad-dsl-grammar},
  while its semantics are in \cref{fig:lrir-semantics}.
In general, the \lrir syntax and semantics 
  are straightforward.
\lrir expressions evaluate to \signal{}s or
  lists of \signal{}s.
Fundamental operations over lists
  (construction, 
  indexing via \texttt{list-ref})
  and \signal{}s
  (\texttt{concat}, \texttt{extract}) 
  are supported.

The only \lrir expression
  which can perform actual computation
  or manipulate state
  (beyond the state merging in \texttt{concat})
  is \texttt{prim}.
\texttt{prim}
  represent instances
  of \textit{hardware primitives} (e.g., LUTs and DSPs).
A \texttt{prim} has a \textit{name}
  (e.g., ``LUT4''), a \textit{configuration},
  and an \textit{input}.

The semantics of a \texttt{prim} expression
  are the semantics extracted from the
  primitive's vendor-supplied Verilog model,
  as introduced in the previous section.
Evaluating \texttt{prim} expressions
  ``drops through'' to these imported
  semantics.
\lrir stores these semantics in
  a context 
  \[\Gamma: \left(\forall p: \textsc{Prim},
            \textsc{Config}\ p\right)
            \rightarrow [\texttt{signal}] 
            \rightarrow [\texttt{signal}]\]
  that takes a configuration
  for a primitive $p$
  and returns a function from
  \texttt{signal}s to \texttt{signal}s
  that captures the semantics
  of $p$ with that configuration.
A configuration of primitive $p$,
  \textsc{Config} $p$,
  is the list of parameters that
  an instantiation of $p$ expects
  (e.g., a Xilinx Ultrascale+ DSP48E2's
  \texttt{AREG}
  and \texttt{BREG}
  parameters, which configure pipelining).
  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
% Subsection: Correctness of the LR Function
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection{Mathematical Model of \lr}
\label{sec:correctness-argument}

\newcommand{\SketchGenFn}{f_{\text{G}\xspace{}}}
\newcommand{\LRSynthFn}{f_{\text{S}\xspace{}}}
\newcommand{\HDLCompFn}{f_{\text{C}\xspace{}}}

Now that we have formalized
  \lr's fundamental components---%
  signals,
   imported Verilog semantics,
   and \lrir---%
  we now formalize the \lr workflow
  as a function \lrfn{}.
We define the function to be
the composition of two sub-functions, which we describe in the below sections:
% \[\lrfn: \Spec \times \Sketch \times \VendorSemantics \rightharpoonup \HDL.\]
$$\lrfn : \Spec \times \Sketch \times \VendorSemantics \times \textsc{Input} \rightharpoonup \HDL = \HDLCompFn \circ \LRSynthFn$$
where $\Spec$ represents hardware designs, $\Sketch$ represents program sketches, $\VendorSemantics$ represents vendor-provided semantics for FPGA components, and $\textsc{Input}$ represents partially symbolic inputs (discussed below).

\lrfn does not include sketch generation,
  instead taking the sketch as an input.
This is for the benefit of stating the correctness properties:
  There is no \textit{a priori} notion of the ``correctness'' of a sketch---%
  a sketch is ``correct'' 
  so long as program synthesis succeeds 
  in filling the holes in the sketch.
A sketch does, however, influence
  the search space for program synthesis, as we note in \hl{todo fill in forward reference}.
  
% This is because our correctness properties
%     state that any synthesized program is correct.
% If an incorrect sketch is generated then synthesis
%     will fail and the correctness
%     properties will hold vacuously.

% \hl{New user input needed: number of clock cycles, among other things?}

% \hl{should sems be gamma?}

%\hl{ben is pointing out that these functions don't actually compose in their current states b/c of their input/output types. we could do a few things: (1) change figure 2 so everything goes through sketch generator. (2) have the two differ but acknowledge the difference. or (3) go full formal and have the semantics "pass through"}

%\hl{see notes under figure. we can just drop fg, and mention that fg isn't necessary to correctness}

% %%%%%  Sketch Generation
%
% NOTE: I've removed this subsection, but we can include it
% later if we think we should. It's not part of the lr function
% so if we do include it, it should be put elsewhere!
%
% \subsubsection{\SketchGeneration}
% \lrfn begins by performing \SketchGeneration
%   on each available sketch 
%   template\footnote{While sketch
%   templates can in theory
%   be provided by users, the \lr tool supplies
%   a set of base templates and, for simplicity
%   of the formalization, we treat
%   them as being internal to the tool.}.
% We formalize \SketchGeneration with the function
% \hl{should also include the spec, b/c we need to know the input/output widths}
%   \[\SketchGenFn :\SketchTemplate \times \ArchDescr \rightharpoonup \Sketch,\]
%   where \(\rightharpoonup\)
%   indicates that this operation is partial.
% $\SketchGenFn$ is partial because a sketch template
%   may use a primitive interface that isn't supported
%   by a given architecture's description.
  
%%%%%  LRSynthesis
\subsubsection{\LRSynthesis: $\LRSynthFn$}
\label{sec:lr-synth-defs}
\lrfn begins by attempting to synthesize an \lrir{} program
  that implements a provided specification using
  a program sketch.
We formalize this component with 
  the $\LRSynthFn{}$ function:
  \[\LRSynthFn: \Sketch \times \Spec \times \VendorSemantics \times \textsc{Input} \rightharpoonup \lrir.\]
$\LRSynthFn$ is partial since
  synthesis may not find an
  implementation for a spec
  using the given sketch---%
  Rosette may conclude that no implementation exists
  or it may time out.
%\hl{Maybe add signal 
  %mention in this section 3.2.2.}
  
\newcommand{\lrinterp}[0]{\mathcal{I}_{Ax}}

We formalize synthesis
  in terms of an interpretation
  function 
  $\lrinterp$,
  which expects
  an \lrir program $p$,
  vendor-provided semantics $e$ 
  (used for interpreting the
  semantics of any \texttt{prim}
  constructs in $p$),
  a vector $\mathbf{i}$ of
  partially symbolic values,
  and a partially symbolic
  state map $m$
  as input values,
  and produces a partially
  symbolic signal as output.
A \textit{partially
  symbolic signal} is a pair $(v, m)$
  of a partially symbolic value $v$
  and a partially symbolic
  state map $m$.
A \textit{partially symbolic value} $v$
  is either a concrete bitvector
  or a symbolic bitvector.
A \textit{partially symbolic map} $m$
  is a map from concrete keys
  to partially symbolic values.
% For a symbolic bitvector $v$,
%   $f(v) = g(v)$ means that $\forall v^\prime, f(v^\prime) = g(v^\prime)$,
%   where $v^\prime$ is a concrete bitvector;
%   that is, we quantify 
%   over all possible instantiations 
%   of the symbolic values.

%\hl{we need to get $(v,m)$/$(v,s)$ consistent across the paper. $s=(v,m)$ is probably better. Needs to be propagated to first half of formalization. -gs}


We define $\lrinterp$ recursively:
\begin{align*}
    \lrinterp(p, e, [\mathbf{v_1}], m) &= \llbracket p \rrbracket (\mathbf{v_1}, m) \\
    \lrinterp(p, e, [\mathbf{v_1}, \mathbf{v_2}, \ldots, \mathbf{v_n}], m)  &= \lrinterp(p, e, \\
    & \hspace{15mm} [\mathbf{v_2}, \ldots, \mathbf{v_n}], \pi_S(\llbracket p \rrbracket (\mathbf{v_1}, m)))
    % a little nasty but it breaks margins otherwise
\end{align*}
% $$\lrinterp(p, e, [\mathbf{v_1}]) = \llbracket p \rrbracket (\mathbf{v_1})$$
% $$\lrinterp(p, e, [\mathbf{v_1}, \mathbf{v_2}, \ldots, \mathbf{v_n}]) = \lrinterp(p, e, [S_{\mathbf{v_1}}(\mathbf{v_2}), \ldots, S_{\mathbf{v_1}}(\mathbf{v_n})])$$
%\hl{i think we can make these prettier -- perhaps use aligned env?}
  where $\llbracket p \rrbracket(\textbf{v})$ evaluates $p$ 
  with respect to its semantics 
  (Figure~\ref{fig:lrir-semantics},
  substituting the elements of $\mathbf{v}$ for each \texttt{signal-var} in $p$), and $\pi_S$ projects
  the state map from a signal.
We similarly define $\mathcal{I}_V(d, \mathbf{i}, m)$,
  which behaves identically to $\lrinterp$,
  except that it takes as input a Verilog design $d$ 
  ($\llbracket d \rrbracket$
  evaluates the design
  according to the Verilog semantics
  in Figure~\ref{fig:verilog-semantics})
  and it does not use $e$,
  since there are no
  \texttt{prim} constructs in Verilog.

Given a hardware design $d \in \Spec$,
  vendor-provided semantics $e \in \VendorSemantics$,
  $s \in \Sketch$,
  and a input vector of partially symbolic values,
  $\LRSynthFn$ is defined as sending the following query to Rosette:
$$\exists p \in \lrir\ \text{s.t.}\ \lrinterp(p, e, \mathbf{i}) = \mathcal{I}_V(d, \mathbf{i}).$$
Synthesis succeeds if Rosette
  finds a program $p$ 
  (constructed by filling the holes in $s$)
  that satisfies this property.
How the input 
  $\mathbf{i}$
  is constructed 
  is an important design decision;
  we describe the specific decisions
  made by \lr
  in 
  \cref{sec:implementation:program-synthesis}.
(Note that our formalization has no need
  to ``privilege'' the clock---%
  it can be treated like any other signal.)

% [[d]]([t]) = [[d]]'(t)
% [[d]](s) = s
% Interp(e, i (vector)) each elem of i = tuple of signals
% Interp(e, [] ::' i) = [[e]](i)
% Interp(e, is ::' i) = let (_, s) = Interp(e, is) in
%   Interp(e, [i.update(s)]
% update: update((sig0, sig1, sig2), s) -> overwrites sig0, sig1, sig2 states with s
% vector i: list of concrete length
% element of vector i: tuple of concrete length
% element of that tuple: signal
% that signal: concrete or symbolic value, 
%   state map has concrete keys, concrete or symbolic values

% symbolic inputs are just variables, always bitvectors
% if Rosette succeeds, then it's quantified for all symbolic values


% The $\LRSynthFn{}$ takes
%   $(d, n) \in \Spec$ and attempts
%   to synthesize a program in \lrir
%   with the semantics of $d$.
% When $d$ is sequential, successful
%   synthesis yields a program that
%   implements $d$ after $n$ clock cycles.
% \hl{Ben: \@Gus I'll need you to give more detail on this}
% $\LRSynthFn{}$ does this by applying the
%   interpreter to the sketch multiple times
%   to simulate each cycle of the clock.
% We can treat a sketch $s$ as a function that
%   outputs values on the wire and 
%   updated state (e.g., incremented clock).
% Then Synthesis tries to find an instantiation
%   $S$ of $s$ such that $S^n(i)$
%   implements $d(i)$.
% \hl{I'm not saying the above correctly, we need
%   fix! But this is the basic idea}.
% Here we use the notation $S^n(i)$ to mean
%   the following:
%   \begin{itemize}
%       \item $S^1(i, \sigma_0)$ is defined to  be
%         the 2-tuple $(v, \sigma_1)$
%         of the output value $v$
%         and state $\sigma_1$
%         resulting from applying
%         $S$ to input $i$
%         with initial state $\sigma_0$.
%       \item $S^{n+1}(i, \sigma_0)$ is defined
%         to be the 2-tuple $(v, \sigma_{n+1})$ 
%         of the output value $v$ and the
%         state $\sigma_{n+1}$ resulting from
%         applying $S$ to input $0$ and state
%         $\sigma_n$.
%   \end{itemize}
%   \hl{Ben: This is not quite what we discussed, but I think it's correct for the formalism. We can go into the more detailed approach w/ multiple interpreter calls in implementation.

% I'm also happy to go into more detail here, but I want to start simple and see what this misses.}
  
% When $d$ is combinational,
%   successful synthesis yields a 
%   program that implements the
%   semantics of $d$ being run a
%   single time (no updated state or
%   clock cycles).


%%%%%  HDL Compilation
\subsubsection{\HDLCompilation}
Next, \lrfn compiles each synthesized design
  to an HDL. This is formalized with the function
  \[\HDLCompFn: \lrir \rightarrow \HDL.\]
This compilation is done as a very 
  straightforward syntactic mapping:
  each component
  of \lrir maps directly to the
  resulting Verilog implementation.

% \subsubsection{Combining Components into \lrfn}
%
% NOTE: We no longer need this section
%
% The \lrfn function is built by combining the three above components, as sketched in the following pseudocode:

% \lstset{style=pystyle}
% \begin{lstlisting}[basicstyle=\normalsize\ttfamily]
% def LR(spec, arch, sems):
%   hdl_impls = {}
%   for t in sketch_templates:
%     sketch = generate_sketch(t, arch)
%     lrir = synth_lrir(spec, sketch, sems)
%     if lrir is not None:
%         hdl_impls.add(compile2hdl(lrir))
%   return hdl_impls
% \end{lstlisting}
% \lstset{style=lispstyle}

  

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
% SubSection: Correctness of LRFN Function
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection{Properties of \lrfn{}}

From the above definitions,
  we can prove certain correctness properties
  about \lrfn{} and $\LRSynthFn$
  if we make certain assumptions about our some components of \lr (our trusted computing base):
\begin{enumerate}
\item \textbf{Rosette}: We assume that Rosette is correct (if it returns a program for a synthesis query, the program satisfies the specification) and sketch-complete modulo timeout (given a sketch, if there exists any solution using that sketch, Rosette will either find a solution or time out---it will not falsely conclude that no solution exists).
\item \textbf{Yosys}: We assume that Yosys's conversion of Verilog into bitvector expressions is correct.
\item \textbf{Vendor-supplied semantics}: We assume that these Verilog files correctly model the semantics of the FPGA components they model.\footnote{
  Note that this assumption may not always be true,
    as we will see in
    \cref{sec:completeness}.
}
\item \textbf{The implementation of $\lrinterp$}: In synthesis, we include an interpreter for the semantics of \lrir. We assume that our implementation correctly implements the claimed semantics in Figure~\ref{fig:lrir-semantics}.
\item $\mathbf{\HDLCompFn}$: We assume that our compilation from \lrir to Verilog is correct and complete. We do not prove this, though it could be done in principle via the semantics in Figures~\ref{fig:lrir-semantics} and~\ref{fig:verilog-semantics}; we rely instead on the simplicity of the implementation.
\end{enumerate}

\subsubsection{Correctness of Synthesis}

We can state this property as follows: 
If $\LRSynthFn(d, s, e, \mathbf{i}) = p$
  for design $d$, 
  sketch $s$,
  vendor-provided semantics $e$,
  and a partially symbolic input $\mathbf{i}$, 
  then for all inputs $\mathbf{i}^\prime$ 
  that are valid concrete instantiations of $\mathbf{i}$,
  $\mathcal{I}_V(d, \mathbf{i}^\prime) = \lrinterp(p, e, \mathbf{i}^\prime)$.
This follows directly from our assumptions about the correctness of Rosette and the correctness of Yosys (which is used to implement $\mathcal{I}_V$) and the definitions in \cref{sec:lr-synth-defs}.

\subsubsection{Sketch-Completeness of Synthesis Modulo Timeout}

Given design $d$,
  sketch $s$,
  vendor-provided semantics $e$,
  and a partially symbolic input $\mathbf{i}$,
  if there exists an instantiation of $s$ called $p$ for which $\mathcal{I}_V(d, \mathbf{i}^\prime) = \lrinterp(p, e, \mathbf{i}^\prime)$ 
  for all
  valid instantiations $\mathbf{i}^\prime$ of $\mathbf{i}$, 
  then $\LRSynthFn(d, s, e, \mathbf{i})$ 
  will either return some value or time out.
This follows from the sketch-completeness (modulo timeout) of Rosette.

\subsubsection{Correctness and Completeness of \lrfn{}}

Given the above properties of $\LRSynthFn$
  and our assumption of the correctness and completeness of $\HDLCompFn$,
  it follows that their composition will have analogous properties:
For design $d$,
  sketch $s$,
  vendor-provided semantics $e$,
  and a partially symbolic input $\mathbf{i}$,
  if $\lrfn(d, s, e, \mathbf{i}) = d^\prime$, then $\mathcal{I}_V(d, \mathbf{i}^\prime) = \mathcal{I}_V(d^\prime, \mathbf{i}^\prime)$ for all valid instantiations $\mathbf{i}^\prime$ of $\mathbf{i}$.
Additionally, with $d$, $s$, $e$, and $i$,
  if there exists some instantiation $p$ of $s$ for which
  $\mathcal{I}_V(d, \mathbf{i}^\prime) = \lrinterp(p, e, \mathbf{i}^\prime)$ 
  for all
  valid instantiations $\mathbf{i}^\prime$ of $\mathbf{i}$,
  then $\lrfn(d, s, e, \mathbf{i})$
  will either return a value or time out.

% \subsection{Correctness of the \lrfn{} Function}

% \hl{Gus:Enumerate the 
%  TCB,
%  if we really want to show
%  that we're serious
%  about correctness Solver, semantics importer (a number of components within Yosys + our $\sim$300 line btor to Racket compiler)}
%  \hl{Ben: The TCB is roughly enumerated after Property 3 is listed.}

% The \lr-Correct property says that
%   for instruction specification $s$, 
%   HDL program $v$,
%   architecture description $d$,
%   vendor-supplied semantics $e$,
%   and inputs $i$,
%   if synthesis and HDL compilation
%   with \lrfn on the instruction
%   specified by $s$
%   and architecture described by $d$
%   results in a HDL implementation $v$,
%   then simulating $v$ on inputs $i$
%   produces the same result as evaluating $s$
%   on inputs $i$. \hl{Should we talk here about how signals aren't relevant to the output comparison, just a way for interpretation to get to the output?}\\

% \newtheorem{property}{Property}
% \begin{property}[\lr-Correct]
% For all \(s \in \Spec\),
%         \(v \in \HDL\),
%         \(d \in \ArchDescr\),
%         \(e \in \VendorSemantics\)
%         and inputs \(\mathbf{i}\)
%   \[\lrfn(s,d,e) = v \implies 
%      \denote{s}_{\text{bv}}(i) 
%          = 
%      \denote{v}_{\text{HDL}}(i).\]
% \end{property}

% \hl{should we have
% (flr(...) = fc(fs(...)) = ...? or just one of those things? in which case, which one?}

% The commutative diagram in \cref{fig:correctness-comm-diagram}
%   shows how the \lr-Correct property
%   is composed from the
%   the Synth-Correct (bottom left triangle) and
%   Compile-Correct (bottom right triangle) properties,
%   which we describe below.
  
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % COMMUTATIVE DIAGRAM FIGURE
% \input{figures/correctness-comm-diagram}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % Synth Correct Theorem Section
% The Synth-Correct property states
%   that when performing \LRSynthesis on a specification produces
%   a result, that result is correct.\\


% \hl{
% TODO:
% we have a big down arrow for what lakeroad program evals to over one timestep, we need a big down arrow that represents time stepping over multiple timesteps.}

% %%%%%%%%%%%%%%%%%%%%%%
% % Synth Correct Theorem
% \begin{property}[Synth-Correct]
%   For all \(k \in \Sketch\),
%   \(s \in \textsc{Spec}\),
%   \(c \in \ArchDescr\),
%   \(e \in \VendorSemantics\),
%   \(p \in \lrir\),
%   and inputs \(\mathbf{i}\),
%   \[\LRSynthFn(k, s, e) = p \implies \denote{s}_{bv}(i) = \denote{p}_{\lrir}(e)(i)\]
% \end{property}
% \label{thm:synth-correct}
% % End Synth Correct Theorem
% %%%%%%%%%%%%%%%%%%%%%%

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % Compile Correct Theorem Section

% The Synth-Correct property asserts that 
%   the \HDLCompilation component preserves semantics.\\

% %%%%%%%%%%%%%%%%%%%%%%
% % Compile Correct Theorem
% \begin{property}[Compile-Correct]
%   For all \(p \in \textsc{Impl}\) and all inputs \(i\),
%   \[\HDLCompFn(p) = v \implies \denote{p}_\lrir(e)(i) = \denote{v}_\HDL(i).\]
% \end{property}
% \label{thm:compile-correct}
% % End Compile Correct Theorem
% %%%%%%%%%%%%%%%%%%%%%%

% The Synth-Correct property depends on
%   the correctness of the synthesis tool (in our case Rosette~\cite{torlak2014lightweight, torlak2013growing}),
%   the correctness of vendor-supplied HDL FPGA primitive semantics,
%   and the correctness of our \lrir interpreter.
% All three of these pieces are real software and are thus subject
%   to bugs in principle,
%   but all are heavily tested and thus reasonable
%   inclusions in our trusted computing base.
% Rosette in particular is a widely-used
%   by the Programming Languages and Program Synthesis communities.
% \lr's core \lrir interpreter is straightforward and
%   lifts bitvector and list manipulation
%   functions from Rosette and Racket.
% As such its correctness depends primarily on that of the host languages, though we  have also extensively tested it.
% All hardware-specific portions of the \lrir interpreter are imported
%   automatically from vendor-supplied HDL primitive specifications.
% We assume that these specs are correct as they are provided as ground truth
%   by vendors.
% Again, we have also validated this 
%   importing process through
%   extensive testing.
% Finally, the Compile-Correct property relies
%   on the simplicity of \lrir: each component
%   of \lrir maps directly to the
%   resulting Verilog implementation,
%   and compilation is a simple
%   syntactic transformation.


  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% GRAVEYARD
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \begin{figure}
% \hl{ben: ltop and ptol should also have some shift functionality}
% \begin{verbatim}
% spec ::= <bv-const> | <bv-var>
%       | <binop> <spec> <spec>
%       | <unop> <spec> <spec>
%       | extract <nat> <nat> <spec>
%       | concat <spec> ...
%       | zero-extend <spec> <nat>
%       | list <spec> ...
%       | list-ref <spec> <nat>

% binop ::= bvand | bvor | bvxor 
%         | bvadd | bvsub | bvmul
        
% unop  ::= bvnot

% impl  ::= bv <bv-const-or-var>
%         | extract <nat> <nat> <impl>
%         | logical-to-physical <ltop> <impl> ...
%         | physical-to-logical <ptol> <impl> ...
%         | primitive <name> <spec> ...

% sketch ::= impl with holes
        
% ltop  ::= bitwise
%         | bitwise-reverse

% ptol  ::= bitwise
%         | bitwise-reverse
% \end{verbatim}
% \caption{Caption}
% \label{fig:syntax}
% \end{figure}
