\chapter*{Acknowledgments}
% \addcontentsline{toc}{chapter}{Acknowledgements}


%% FUNDERS

% Before acknowledging the people
%   who made this happen,
%   let me first acknowledge 
%   the SRC CRISP and ADA centers,
%   the DARPA RTML project,
%   Intel,
%   the U.S. Department
%   of Energy,
%   and the NSF,
%   who paid my stipend.

%% PEOPLE

First, allow me to thank
  everyone in my life 
  \textit{not} mentioned in the following paragraphs.
My PhD was far more than just a collection
  of projects;
  for the past six years,
  it was my life.
Anyone who has contributed to my life
  has contributed to
  this dissertation.


Thanks to my early research mentors.
To Vijay Narayanan and Jack Sampson at Penn State.
Without their early mentorship,
  I would have never considered
  the PhD a viable option.
To my initial research mentors in SAMPL:
  Thierry Moreau, Tianqi Chen, Jared Roesch, and Luis Vega.
Thank you for bringing me on
  to TVM
  and introducing me to compilers research.
To my advisor Luis Ceze,
  who hired me on at UW
  and changed
  the course of my entire life.
I could never thank Luis enough
  for the impact that single decision
  had on me.

  
Thanks to everyone I've worked with in industry.
To Aart Bik, Mangpo Phothilimthana,
  and Penporn Koanantakool
  at Google
  for their mentorship throughout my 
  extended 2021 internship.
To Aart and Penporn
  for bringing me onto the MLIR team,
  and to Mangpo for her patient tutelage
  in applying machine learning to compilers.
To Noah Evans and everyone at Sandia,
  for bringing me on
  and being interested in my research.
To Nina, Jannis, Claire, and the entire YosysHQ team
  for their support.
To Jin Yang and Jeremy Casas at Intel.

Thanks to every research team
  I've worked with.
To the Glenside team: Andrew Liu, Steven Lyubomirsky, Scott Davidson, Joseph McMahan, Michael Taylor, Zach Tatlock, and Luis Ceze.
To the 3LA team: Bo-Yuan Huang, Steven, Yi Li, Mike He, Thierry Tambe, Akash Gaonkar, Vishal Canumalla, Andrew Cheung, Gu-Yeon Wei, Aarti Gupta, Zach, and Sharad Malik.
To the Lakeroad team: Ben Kushigian, Vishal, Andrew, Steven, Sorawee Porncharoenwase, Ren\'e Just, Gilbert Bernstein, and Zach.
To the Churchroad team:
  Dan Petrisko,
  Colin Knizek,
  Chandra Nandi, Zach,
  Jonathan Balkind, Zach Sisco and Thanawat Techaumnuaiwit.


Thanks to those who volunteered to be on my committee---%
  to Mike Taylor,  Scott Hauck, 
  and Gilbert Bernstein.
Thanks especially to Gilbert,
  who invested time and effort
  into making me a sharper, more precise
  PL researcher
  in my last year at UW.
Gilbert was brave enough to
  attempt to teach me about coinduction---%
  and persistent enough to (mostly) succeed.
I will always appreciate his willingness to
  take the time and effort
  required
  to explain difficult concepts.

Thanks to the MISL lab
  for letting me hang around
  all these years,
  despite my utter lack of any formal
  training in biology.
Thanks to Chris Thachuk
  and Lancelot Wathieu
  for taking the
  Fridge Compiler
  from half-working course project
  to published paper.
Thanks to Jeff Nivala,
  Gwen Roote, Carina Imburgia,
  and the whole
  FPCA team
  for letting me tag along
  on the project.
A more down-to-earth lab I may never find.

To the entire Cornell Capra lab.
To Rachit Nigam, who
  served as my model of 
  success
  in programming languages and hardware design research.
To Adrian Sampson,
  who I can always look to for advice.
To Priya Srikumar,
  whose brief presence in PLSE
  in the summer of 2023
  was a blessing.
Priya's passing was felt
  throughout the Programming Languages
  research world,
  into their communities in Ithaca and Seattle,
  and beyond.
I will always remember Priya sharing
  a Tom Wayman poem with me,
  titled
  ``Did I Miss Anything?'',
  an excerpt of which
  will always remind me of them:
\begin{quote}
Nothing. When you are not present\\
how could something significant occur?
\end{quote}
% My last messages with Priya were from the day they left Seattle. They wanted to say goodbye in person, but I was out of town. The last message they sent me was, “aww no worries!! im leaving tomorrow and i wanted to say bye, but we’ll see each other again!!” Someday, somewhere, maybe we will.

\clearpage
To the PLSE lab.
It is harder to imagine a
  better lab
  exists anywhere on the planet.
The energy of the lab
  can be felt
  even before stepping through the door.
Whether it's laughter,
  or spirited research discussion,
  or someone playing one of the lab's many instruments,
  or even just the peaceful silence
  of face-in-laptop
  research progress,
  the energy of the lab is one of a kind.
Without realizing it,
  the lab has become my family;
  leaving it
  has been one of the most challenging things
  I've ever done.
I can only hope to start something
  half
  as vibrant, half as connected,
  half as supportive,
  wherever I end up.

Thanks to my mentee Hannah Leung,
  whose unyielding positivity
  and bravery
  in the face of challenges
  the likes of which I'll never 
  myself experience
  was a constant source of inspiration.
Though she may not know it,
  I've learned more from her
  than she has from me.
  

To my mentees
  Vishal Canumalla
  and Andrew Cheung.
Thanks for trusting me
  to mentor you in research and in life,
  and thanks for your friendship
  along the way.
I am so proud of you guys,
  and cannot wait to see what you do next.
Furthermore, thanks to Vishal's family,
  the Canumallas,
  who brought me in.
Thank you Anu and Sridhar
  for making me feel at home
  and making sure I'm well fed.

Thanks to Race Condition Running
  for the miles and miles of fun
  over the years.
Through my best and worst times
  in the PhD and in my personal life,
  the Saturday run and brunch
  were there for me
  as a consistent source of companionship
  and connection.
Thanks to Zach, Nick Walker, Ellis,
  Chandra, Max,
  and everyone else who
  kept the club going.
Thanks to everyone else who kept me active
  during my PhD---%
  thanks to Eric Zeng and Chien-Yu Lin
  for getting me out in the backcountry
  when the snow was good
  and to Gwen Roote for keeping me humble
  on the BJJ mats.
  
Thanks to the lab jam crew---%
  Ben, Carina, Rohan, and everyone else---%
  for providing a safe space
  to be a musical beginner.



To Ben Kushigian---a wonderful coauthor and friend.
  
To Nick and Danielle.
To Holly and Peter.
To Jared.

To Anjali, for all of the time you've given me.

To \texttt{\#binary-blobs}, \texttt{\#spicelords}, and associated friends:
  Pratyush Patel, Chien-Yu Lin,
  Jacob Van Geffen,
  Dan Petrisko, 
  Luzdary Ruelas,
  Ellis Michael,
  Max Ruttenberg,
  and Katie Lim.
Thanks to the countless hours
  of Twilight Imperium.


To Max Willsey and Sami Davies.
Without Max's urging,
  I may have never joined UW;
  even after I joined UW, without Max's urging,
  I may have never joined PLSE.
Max's quiet, confident intelligence
  served as a model for the entire PLSE lab.
I was lucky to witness
  a golden age of PLSE research,
  with Max's \texttt{egg} at the forefront.
Max showed me that a new project,
  an empty buffer,
  a clean slate,
  was nothing to be scared of
  if you face it with confidence and excitement.
And to Sami,
  whose intimidating intelligence
  is softened by her infectious smile
  and sincere desire for connection.


To Steven Lyubomirsky---%
  \textit{Lex Lyubomiricus} will forever guide my research.
Thank you for your constant mentorship
  at every stage of my research career.
Steven helped me present my first Programming Languages
  Reading Group paper in 2018---%
  ``Build Systems \`a la Carte'' by Mokhov et al.---%
  before I had any
  formal Programming Languages research
  under my belt.
Six years later, in 2024,
  Steven helped me publish the final
  first-author publication of my PhD,
  the Lakeroad paper.
  
To Chandra Nandi---%
  I could not have asked for a better
  research mentor and friend.
Chandra helped
  in every single one of my paper pushes
  in grad school,
  simply out of the love
  of camaraderie.
She taught me how
  to feel the unabashed joy
  that comes 
  from working hard
  on cool problems
  with your closest friends.


To Jeff's sons,
  Christian, Nick, and Aaron.
You all are so much like your father.
Thanks for 
  bucking the Seattle Freeze
  and making me feel at home here
  in my first 
  few years.
Here's to 
  many, many more rounds
  of pitch and putt.

To Eric, Rick, Zach,
  and the other men in the Montlake Men's Group.
Your support
  and mentorship
  helped me navigate the most challenging chapters
  of my life to date.
  
To my friends
  from the previous chapters of my life
  that I've been lucky enough
  to bring into this one:
  Spencer Norris, Imaz Athar,
  Anthony Marucci, Tim Lagnese, Jacob Brunette,
  and Drew Abbott.
Thank you to my oldest 
  and best friends,
  Spencer and Imaz.
There is something
  bittersweet about how,
  more than almost anything else,
  it's simply \textit{time} 
  that is the key ingredient
  for a deep relationship---%
  I'm so lucky to have invested
  my time in you guys.
You two are
  and will always be
  my brothers.
It has been an honor and a joy
  to grow with you
  through life's changes,
  to have you guys help me
  through my lowest lows
  and celebrate with me during my highest highs.
HSNE, boys.
Thanks to
  Anthony, Tim, and Jacob.
I could not have picked a better crew
  of friends to move to Seattle with,
  to spend my 20s with.
May our caps always fly true.
To Drew and Em---%
  we've rarely lived in the same city,
  but you two are family to me.
  
Boundless love and thanks to my older brother Simon.
In all my 29 years,
  he has never
  been far from my mind.


Thank you to Michaela,
  whose five years of love and loyalty transformed me
  into the man I am.
Her support carried me through the single
  continuous paper push
  that was the two years
  of my PhD.
Her patience with the all-consuming nature
  of my PhD
  was more than I could have asked for.
Her genuine care
  for everyone around her
  showed me that there is nothing more to life
  than the people in it.
There is no one more deserving
  of happiness.
Furthermore, thanks to the entire Brock and Lowe family,
  who brought me in
  without question,
  who made me a part of their family.
To Lisa Lowe, whose unhesitating love
  from the moment we met
  taught me the meaning of
  ``unconditional''.
To Ray Brock,
  who was
  a role model for men
  in a world where they are severely lacking.
I am immensely lucky to have known him.
It is through the Lowes
  that I came to understand
  the meaning of family---%
  a gift I could never
  repay.

  
Finally, to my advisor, Zach Tatlock.
Zach is the kind of researcher
  all should strive to be.
Intelligent, hardworking,
  and selfless to a fault.
Zach brought me on
  when he didn't need to.
He has a steady supply
  of top-tier PL students.
Bringing me on, as
  a student with zero formal programming languages
  research experience,
  was in many ways a questionable decision.
But, like Luis's decision to admit me,
  that single decision altered the course
  of my entire life.
Everything I know of research
  I learned first from Zach.
But Zach is far more than just a good researcher.
Zach is, in fact,
  the only thing I really strive to be---%
  a good man.
I've had very few role models in life,
  and Zach is first among them.
Thank you, Zach---%
  none of this could have happened
  without you.